\documentclass{jot}

\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{microtype} % optional, for aesthetics
\usepackage{tabularx} % nice to have
\usepackage{booktabs} % necessary for style

%\usepackage[bookmarks,bookmarksopen,bookmarksdepth=2]{hyperref} %sections as bookmarks in adobe pdf
%\hypersetup{pdfpagemode=UseNone}


%%% Article metadata
\title{Domain-Specific Model Distance Measures}
\runningtitle{Domain-specific distance measures}

\author[affiliation=UL, nowrap] % , photo=FILE]
    {Manuel Wimmer}
    {is a ... in ... at ....
    Contact him at \email{EMAIL}, or visit \url{URL}.}

\author[affiliation=UdeM, nowrap] % , photo=FILE]
{Eugene Syriani}
{is an associate professor in the department of computer science and operations research at Universit{\'e} de Montr{\'e}al.
	Contact him at \email{syriani@iro.umontreal.ca}, or visit \url{www.iro.umontreal.ca/~syriani}.}

\author[affiliation=UV, nowrap] % , photo=FILE]
{Robert Bill}
{is a ... in ... at ....
	Contact him at \email{EMAIL}, or visit \url{URL}.}

\affiliation{UdeM}{Universit{\'e} de Montr{\'e}al}
\affiliation{UV}{University of Vienna}
\affiliation{UL}{University of Linz}

\runningauthor{Wimmer et al.}

\jotdetails{
    volume=V,
    number=N,
    articleno=M,
    year=2011,
    doisuffix=jot.201Y.VV.N.aN,
    license=ccbynd % choose from ccby, ccbynd, ccbyncnd
}

% Put edit comments in a really ugly standout display
\usepackage{ifthen}
\usepackage{amssymb}
\newboolean{showcomments}
\setboolean{showcomments}{true} % toggle to show or hide comments
\ifthenelse{\boolean{showcomments}}
{\newcommand{\nb}[2]{
		\fcolorbox{gray}{yellow}{\bfseries\sffamily\scriptsize#1}
		{$\blacktriangleright$#2$\blacktriangleleft$}
	}
	\newcommand{\version}{\emph{\scriptsize$-$working$-$}}
}
{\newcommand{\nb}[2]{}
	\newcommand{\version}{}
}

\newcommand\es[1]{\nb{ES}{\textcolor{red}{\textsl{#1}}}}
\newcommand\mw[1]{\nb{MW}{\textcolor{red}{\textsl{#1}}}}
\newcommand\rb[1]{\nb{RB}{\textcolor{red}{\textsl{#1}}}}

% For URLs in the references
\usepackage{url}
\usepackage{hyperref}
%\usepackage{verbatim}

% Advanced Math
%\usepackage{amsmath}
%\usepackage{amssymb}	% double line font letters (for number sets N,Z,D,Q,R)

% Images and Floats
\usepackage{epsfig}
%\usepackage{epstopdf}

%\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{lscape}			% landscape
%\usepackage{array}			% align vertically in tables
%\usepackage{multirow}
%\usepackage{rotating}

% Theorems, Definitions, ... (always after 'amsmath')
%\usepackage{amsthm}
%\theoremstyle{definition}
%\newtheorem{prop}{Property}%[section]
%\newtheorem{defn}{Definition}%[section]

% Algorithm environment
%\usepackage{algorithm}
%\usepackage{algorithmic}

% Commutative diagrams
%\usepackage{diagrams}

%\usepackage{enumitem}

%\usepackage{listings}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Standard shortcuts
\newcommand{\eg}{e.g.,~}										% exempli gratia (for the sake of example)
\newcommand{\ie}{i.e.,~}										% id est (that is)
\newcommand{\etal}{~et al.}									% et alia (and others)
\newcommand{\Fig}[1]{Figure~\ref{#1}}  			% choose Fig. or Figure, depending on the style
\newcommand{\Table}[1]{Table~\ref{#1}}	    % Table reference
\newcommand{\Sect}[1]{Section~\ref{#1}}	  	% section name always with a capital S
\newcommand{\Model}[1]{\textsf{\small{#1}}} % name of any modeling artifact (e.g., formalism, model element, rule, ...)
\newcommand{\Code}[1]{\texttt{\small{#1}}}	% inline code
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}	% scientific notation: x.10^y

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

\begin{abstract}
A lot of research was invested in the last decade to develop differencing methods for models to identify the changes performed between two model versions. 
A difference model captures these changes. However, less attention was paid to distance computations of model versions. While different versions of a model may have a similar amount of differences,
the distance to a base model may be drastically different. Therefore, we present in this paper distance metrics for models, a method to automatically generate tool support for computing distance measures and show the benefits of distance measures over differences in searching for model evolution explanations. The results of running different experiments show...
\end{abstract}

\keywords{Three; Keywords.}

\section{INTRODUCTION}
\input{01_intro.tex}

\section{BACKGROUND}
\input{02_background.tex}

\section{DOMAIN-SPECIFIC DISTANCES}
\input{03_approach.tex}

\section{EVALUATION AND DISCUSSION}
\input{04_eval_discussion.tex}

\section{RELATED WORK}
\input{05_related.tex}

\section{CONCLUSION}
\input{06_conclusions.tex}

\bibliographystyle{alphaurl}
\bibliography{biblio}

\abouttheauthors

\begin{acknowledgments}
This work has been sponsored by
\end{acknowledgments}

\end{document}
